---
title: "Merging: Exercise"
---

Now we have gotten to know our `characters` data set a bit more. However, the personality ratings are not included yet. 
For that, we need to load the `psych_stats` data set as well, which contains the ratings for all characters. We have already done that in [Exercise 2 of the Load data chapter](https://nickhaf.github.io/r_tutorial/qmd/load_data/load_data_exercise.html#exercies-2), so here is the code again: 

```{r load characters, echo = FALSE}
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
```

::: {.calout-caution icon="false"}
```{r load psych_stats}
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)
```
:::

Now we want to combine both data frames into one, so all information belonging together can be found in the same data frame. 

## Exercise 1
Merge the `characters` data frame and the `psych_stats` data frame on a common column.

::: {.callout-tip collapse="true"}
## Hint
Identify the common columns. Are they named the same in both data frames? Look at the documentation of `?merge` to see, how you can merge data frames that don't have identically named columns. 
:::

::: {.callout-caution collapse="true"}

## Solution
First, let`s take a look at both data sets again:
```{r}
str(characters)
str(psych_stats)
```

It seems like both data frames have a column containing an ID for the character. We can use that column for merging:
```{r}
characters_stats <- merge(
  x = characters,
  y = psych_stats,
  by.x = "id", 
  by.y = "char_id"
)

str(characters_stats)
```
Worked like a charm!

:::
