---
title: "Merging: Exercise"
---

::: {.callout-caution icon="false" collapse="true"}
## Previous code

```{r, message = FALSE}
# install.packages("tidyverse")
# install.packages("here")

library(tidyverse)
library(here)

## Load the data
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

## Reshape into long format:
psych_stats <- psych_stats %>%
  pivot_longer(cols = messy_neat:innocent_jaded, 
               names_to = "question", 
               values_to = "rating")

## Take a look at the data sets
str(characters)
str(psych_stats)
```
:::

Now we have gotten to know our `characters` data set a bit more. However, the personality ratings are not included yet. For that, we need to combine it with the `psych_stats` data set.

## Exercise 1

Merge the `characters` data frame and the `psych_stats` data frame on a common column.

::: {.callout-tip collapse="true"}
## Hint

Identify the common columns. Are they named the same in both data frames? Look at the documentation of `?merge` to see, how you can merge data frames that don't have identically named columns.
:::

::: {.callout-caution collapse="true"}
## Solution

First, let's take a look at both data sets again:

```{r}
str(characters)
str(psych_stats)
```

It seems like both data frames have a column containing an ID for the character. We can use that column for merging:

```{r}
characters_stats <- merge(
  x = characters,
  y = psych_stats,
  by.x = "id", 
  by.y = "char_id"
)

str(characters_stats)
```

Worked like a charm!
:::
