---
title: "Merging: Exercise"
---



::: {.callout-caution icon="false" collapse="true"}
## Previous code

```{r, message = FALSE}
library(tidyverse)

## Load the data
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

## Take a look at the data sets
str(characters)
str(psych_stats)
```
:::

Now we have gotten to know our `characters` data set a bit more. However, the personality ratings are not included yet. For that, we need to load the `psych_stats` data set as well, which contains the ratings for all characters. 
Now we want to combine both data frames into one, so all information belonging together can be found in the same data frame.

## Exercise 1

Merge the `characters` data frame and the `psych_stats` data frame on a common column.

::: {.callout-tip collapse="true"}
## Hint

Identify the common columns. Are they named the same in both data frames? Look at the documentation of `?merge` to see, how you can merge data frames that don't have identically named columns.
:::

::: {.callout-caution collapse="true"}
## Solution

First, let\`s take a look at both data sets again:

```{r}
str(characters)
str(psych_stats)
```

It seems like both data frames have a column containing an ID for the character. We can use that column for merging:

```{r}
characters_stats <- merge(
  x = characters,
  y = psych_stats,
  by.x = "id", 
  by.y = "char_id"
)

str(characters_stats)
```

Worked like a charm!
:::
