# Data structures

Please take a look at the different [data types](data_types.qmd) that exist in R, if you are not familiar with them, before continuing with this chapter.

There are five main data structures in R which differ on their dimensions (one dimension, two dimensions, n dimensions) and the type of the elements they are containing (same type, different types):[^structures_table]

```{r, echo = FALSE, results = 'asis'}
dat_structs <- data.frame(" " = c("1d", "2d", "nd"),
                          Homogeneous = c("atomic vector", "matrix", "array"), 
                          Heterogeneous = c("list", "data.frame", "")
)

knitr::kable(dat_structs, "pipe",
             col.names = c("", "Homogeneous", "Heterogeneous")
             ) 

```

[^structures_table]: Table from [Advanced R](http://adv-r.had.co.nz/Data-structures.html).

Let's take a closer look at each one of them:

## Vectors
Atomic vectors (from hereon only called vectors) contain elements of only the same type: 

```{r}
num_vec <- c(2023, 8, 8)

char_vec <- c("This", "is", "a", "vec", ".")

log_vec <- c(TRUE, FALSE)

```

The `c()` stands for combine, or concatenate, and is the basic function for building a vector out of single elements. 


# Exercises
1. Load the `tutoR` package and start the exercises about `data structures` by typing:

```{r}

```

