## What is R?

## 

## Installation

## R

## RStudio

### Structure of the RStudio interface

#### Script Pane

#### Console {#console}

```{r}
"Hello World!"
```

Tadaa, we just wrote our first line of code.

## First Steps

Some of the most basic things you can do in `R` is are basic mathematical operations. In your [Console](#Console), type the following code lines and press `Enter`.

```{r, eval = FALSE}
# sum
1 + 2

# substract
2 - 4

# multiply
4 * 4

# divide
3/2
```

By the way: Code lines, that are preceded by a `#` are commented out, and will not be evaluated.

## Workflow

When only coding in our [Console](#Console) we quickly run into the question where to save our code for further evaluation. Therefore, we want to write all our Code that is worth saving into a [Skript](#Skript%20Pane).

Over time, it will become increasingly hard to organize all your files, working directories and workspaces in a reasonable manner. For this, RStudio Projects can be used (highly recommended).

```{r, exercise = TRUE}
# 1. Create a new folder folder for this tutorial, where all your files will go.
# 2. Create a new RStudio project and open it. 
# 3. Create a new R Script and open it. 
```
