# What is R?

## 

# Installation

## R

## RStudio

### Structure of the RStudio interface

#### Script Pane

#### Console

```{r}
"Hello World!"
```

Tadaa, we just wrote our first line of code.

### RStudio Cloud

# First Steps

Some of the most basic things you can do in `R` is are basic mathematical operations. In your [console](#console), type the following code lines and press `Enter`.

```{r, eval = FALSE}
# sum
1 + 2

# substract
2 - 4

# multiply
4 * 4

# divide
3/2
```

By the way: Code lines, that are preceded by a `#` are commented out, and will not be evaluated.

## Workflow

### Scripts

When only coding in our [console](#console) we quickly run into the question where to save our code for further evaluation. Therefore, we want to write all our code that is worth saving into a [script](#script_pane).

### RStudio Projects

Over time, it will become increasingly hard to organize all your files, working directories and workspaces in a reasonable manner. For this, RStudio Projects can be used (highly recommended).

### Exercises

1.  Create a new folder folder for this tutorial, where all your files will go.
2.  Create a new RStudio project and open it.
3.  Create a new R Script and open it.

## Packages

### Package for this tutorial

There is a package available for this tutorial, which will provide interactive exercises. Please install:
