# New columns

Adding new columns to a data frame is pretty straight forward. We just define the column name, and then assign it some input:

```{r}
persons <- data.frame(name = c("Anna", "Alex", "John", "Jessi"),
                      age = c(19, 17, 18, 18),
                      birth_month = c("Jan", "Sep", "Oct", "Mar"),
                      big5_extro = c(3.5, 2, 4.5, 4.2)
                      )

persons$big5_neuro <- c(3, 4, 1, 2)

```

Or, using the `tidyverse`:

::: tidy
```{r, message = FALSE}
library(dplyr)
persons <- persons %>%
  mutate(big5_open = c(3, 4, 3, 4))
```
:::

# ifelse

We can also add new values conditionally, for example by using the `ifelse()` function:

```{r}
persons$big5_neuro_dichotomous <- ifelse(test = persons$big5_neuro >= 5/2,
                                         yes = "high",
                                         no = "low")
persons
```

::: tidy
```{r}
persons <- persons %>%
  mutate(big5_extra_dichotomous = ifelse(test = persons$big5_neuro >= 5/2,
                                         yes = "high",
                                         no = "low")
  )
persons
```
:::

# Merging

```{r}
times <- readRDS(file = "./raw_data/times.rds")

head(times)
```

Hmm, looks like we can connect our data sets using the `ID` column:

## Merging data frames

# missings

Finally, we have to talk about missing values. Missing values are denoted in R by `NA` (or `NaN` in some cases). We have to deal with them
