# Setup

## R

## RStudio

### Optional: posit Cloud

In case you don't use a notebook where you can install R and RStudio, or you don't want to, you can use the posit Cloud service. It can be run in you browser, and provides the same functions and interface as if you were working with your own RStudio installation. And it's free as well (at least 25 hours per month, but that should be more than enough for this workshop and probably for the following week).

1.  Go to [posit Cloud](https://posit.cloud/).
2.  Click on `GET STARTED`.
3.  Create an account (it's free) and login.
4.  On the upper right, click on `New Project`. This will create a new RStudio project, which you can use the same way as described in this tutorial.

### Structure of the RStudio interface

#### Script Pane

```{r}
print("Hello World!")
```

Tadaa, we just wrote our first line of code.

#### Console {#console}

You can also work directly in the console, for example to try out things. Type, into your console:

```{r}
# Sum two values
10 + 5
```

By the way: Code lines, that are preceded by a `#` are commented out, and will not be evaluated.
