---
title: "Reshaping: Exercise"
---

::: {.callout-caution icon="false" collapse="true"}
## Previous code

```{r, message = FALSE}
# install.packages("tidyverse")
# install.packages("here")

library(tidyverse)
library(here)

## Load the data
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

## Take a look at the data sets
str(characters)
str(psych_stats)
```
:::

## Exercise 1

Take a look at the data frame `psych_stats`. Which format does it have?

-   [ ] Wide format
-   [ ] Long format
-   [ ] None of the above

::: {.callout-caution collapse="true"}
## Solution

-   [x] Wide format
-   [ ] Long format
-   [ ] None of the above

Each unit of observation, in this case each character, only has one row.
:::

## Exercise 2

Reshape it, so there are only three columns in the data set: `char_id`, `question` and `rating`.

::: {.callout-tip collapse="true"}
## Hint

You can select multiple columns like this: `column_1:column_10`.
:::

::: {.callout-caution collapse="true"}
## Solution

```{r}
psych_stats <- psych_stats %>%
  pivot_longer(cols = messy_neat:innocent_jaded, 
               names_to = "question", 
               values_to = "rating")

head(psych_stats)

```

Now we have multiple rows for every character, but all question ratings are nicely aligned in one column.
:::

## Exercise 3

Try to reshape the data into long format again.

::: {.callout-caution collapse="true"}
## Solution

```{r}
psych_stats %>%
  pivot_wider(id_cols = char_id, 
               names_from = "question", 
               values_from = "rating")

```

This is how we got it! But scratch that, it was just for the sake of the exercise. We want to use `psych_stats` in the long format from now on.
:::
