---
title: "Wide and Long Format"
---

In this chapter, we will look at a simpler data set that makes it a bit easier to explain reshaping as our relatively complex `athletes` data set, which already has a sensible shape.

Let's define our own data set:

```{r}
inhabitants_wide <- data.frame(
  country = c("China", "India", "USA"),
  inhabitants_2021 = c(1425893465, 1407563842, NA),
  inhabitants_2022 = c(1425857720, 1420939232, 338903174)
)
```

## Wide Format

In the wide format, we have only one row per unit of analysis. In this example, each country has it's own row:

```{r}
head(inhabitants_wide)
```

However, the variable `inhabitatns` is stretched over multiple rows. In many cases it makes sense to reshape the data, so the `inhabitants` are all put into one column:

## Long Format

This would be the long data format, where each unit of analysis is spread over multiple rows.

## Reshaping from Wide to Long Format

To get from Wide Format to Long Format we can use the `pivot_longer()` function from the `tidyverse`:

::: tidy
```{r}
inhabitants_long <- inhabitants_wide %>%
  pivot_longer(
    cols = c("inhabitants_2022", "inhabitants_2021"),
    names_to = "year",
    values_to = "inhabitants"
  )

head(inhabitants_long)
```
:::

## Reshaping from Long to Wide Format

In other cases it might happen, that multiple variables are put into the same column, togehter with an identifier column:

```{r, echo = FALSE}
inhabitants_long_2 <- inhabitants_wide %>%
  mutate(inhabitants_2021 = c(9597000, 3287000, 9834000)) %>%
  rename(area = inhabitants_2021) %>%
  pivot_longer(!country, names_to = "variable", values_to = "value") %>%
  data.frame()
```

```{r}
inhabitants_long_2
```

In that case it can make sense to spread the the distinct variables into two columns:

```{r}
inhabitants_wide_2 <- inhabitants_long_2 %>%
  pivot_wider(
    id_cols = "country",
    names_from = "variable",
    values_from = "value"
  )

inhabitants_wide_2
```
