---
title: "Plotting"
---

::: callout-note
This chapter is optional.
:::

::: {.callout-caution collapse="true"}
## Previous code

::: tidy
```{r, message=FALSE}
library(tidyverse)

athletes <- readRDS(file = here::here("raw_data", "athletes.rds"))
world_coordinates <- readRDS(file = here::here("raw_data", "world_coordinates.rds"))

medal_counts <- athletes %>%
  filter(Medal == "Gold") %>%
  group_by(Region) %>%
  count(Medal)

medal_countries <- world_coordinates %>%
  rename("Region" = region) %>%
  left_join(medal_counts)
```
:::
:::

We now want to take a closer look at how `ggplot2` works. We already had a quick glimpse at it: Plots are build from different layers to create complex output. There are endless possibilities for different plot types, look at the [R graph gallery](https://r-graph-gallery.com/) for some inspiration and code.

First, let's plot a relatively simple plot to get you familiar with how `ggplot2` works. After that, we will use the preparation we have done in the last chapters to plot the number of gold medals each country has won over the years on a world map, which gets slightly more complex.

# First plot

For this plot, we want to find the country with the most medal winners for each sport. The preparation is pretty similar to what we have done before:

::: tidy
```{r}
best_by_sport <- athletes %>%
  ## Get all gold medalists
  filter(Medal == "Gold") %>%
  ## Group them by sport and region
  group_by(Sport, Region) %>%
  ## count the number of medals each country has per sport category
  count(Medal) %>%
  ## Now only group by sport, so we can extract the maximum medal row by sport, and not by sport and country
  group_by(Sport) %>%
  ## Extract the country with the most medals
  slice(which.max(n))

head(best_by_sport)
```
:::

## `ggplot()`

In general, a `ggplot` starts with the `ggplot()` function. In it we define the data we want to use, and some aesthetics. The `ggplot()` function then draws our (currently still empty) plotting area, with the defined axes (see next section).

## `aes()`

Aesthetics set parameters **dependent on the data**. In most cases, we will define our x and y axis here. We can also group data together by groups found in a column. If we want the data to have a different color, form, filling etc. depending on values in a column, we can define that here as well (we will look at that later on).

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- ggplot(
  data = best_by_sport,
  aes(
    x = Sport,
    y = n
  )
)

p
```
:::

In this case, the sport is plotted on the x axis and the number of gold medals (`n`) on the y axis.

## `geom_()`

The geoms do the actual plotting. For example, if we want a barplot:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p +
  geom_col()
```
:::

Looking pretty boring, right? Let's give each country another color by defining the `fill` aesthetic. Also, lets order the x axis depending on the number of gold medalists:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  geom_col(aes(fill = Region, x = reorder(Sport, n)))
p
```
:::

Correct! We can define the aesthetics also within the `geom_()` functions. In this case they will only be used for that specific function, and not for the whole plot (which would be the case if we had defined the `fill` aesthetic in the `ggplot()` function):

## Global vs local `aes()`

To illustrate this, we use the `colour` aesthetic instead of `fill`:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p_2 <- p +
  geom_col(aes(colour = Region, x = reorder(Sport, n)))
p_2
```
:::

Now let's add some points:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p_2 +
  geom_point()
```
:::

This gives us black points now. But if we define our aesthetics within the `ggplot()` function, the points have the same colors as the bars, as now the aesthetics apply to all layers:
```{r}
#| fig-height: 8
#| fig-width: 11

ggplot(
  data = best_by_sport,
  aes(
    x = Sport,
    y = n,
    colour = Region
  )
) +
  geom_col() +
  geom_point()
```


Also note that we layer another geom (`geom_point()`) over our plot (that's what I meant earlier on by saying plots consist of different layers).

## General plotting options

We can tweak all aspects of the appearance of a plot. For example, we might want to turn the x axis labels by 90 degrees to actually make them readable:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p
```
:::

Or we could label the bars with the country:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  geom_text(aes(label = Region), hjust = -0.3, angle = 90, size = 2.5)
p
```
:::

Or use different colors and a different theme:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

# install.packages("viridisLite")

p <- p +
  theme_classic() +
  ## And turn the axis labels again, because the new theme has overwritten our theme
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ## Specify which colors are used for the filling. They are from the package viridsLite, so you might need to install it.
  scale_fill_manual(values = viridisLite::viridis(19))

p
```
:::

Finally, change the title and axis labels:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p +
  ggtitle("Country with the most Olympic gold medal winners by sport") +
  xlab("Sport") +
  ylab("Number of gold medal winners")
```
:::

::: callout-tip
Of course we don't have to assign every intermediate step to the `p` object. Normally, we would just combine all layers by using the `+`.
:::

Pretty cool! Now we know that the most Olympic Tug-Of-War gold medalists are from the UK! Also note that we are looking at the number of people from each country winning a gold medal, so team sports are counted multiple times.

# Second plot

Let's take a look at a different plot, just to see what can be easily achieved with `ggplot2` (and we have prepared the data to do that, so we don't want to waste the work).

So, let's build our coordinate system:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- ggplot(
  data = medal_countries,
  mapping = aes(x = long, y = lat, group = group)
)
p
```
:::

Now we add the map by using the polygon geom:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  geom_polygon(aes(fill = n))
p
```
:::

In this case the polygon filling (which are our countries) will change depending on `n` (the number of gold medalists).

In the end we can clean up the plot by rescaling the x and y axis, provide a new colour scale, titles and theme:

::: tidy
```{r}
#| fig-height: 8
#| fig-width: 11


p +
  ## Rescale the plot
  coord_fixed(1.3) +
  ## Use another colour palette:
  scale_fill_distiller(palette = "YlOrBr", direction = 1) +
  ## Plot title
  ggtitle("Olympic Gold Medal winners") +
  ## Legend title:
  guides(fill = guide_legend(title = "Gold medal \n winners")) +
  ## Minimal theme:
  theme_void()
```
:::

Note that the medals for team sports are counted multiple times, as our data contained the gold medal wins by person, and not by discipline.
