---
title: "Plotting: Exercises"
---

::: callout-note
These exercises are optional.
:::

::: {.callout-caution icon="false" collapse="true"}
## Previous code

```{r, message = FALSE}
# install.packages("tidyverse")
# install.packages("here")

library(tidyverse)
library(here)

## Load the data
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

## Reshape into long format:
psych_stats <- psych_stats %>%
  pivot_longer(
    cols = messy_neat:innocent_jaded,
    names_to = "question",
    values_to = "rating"
  )

## Merge it
characters_stats <- merge(
  x = characters,
  y = psych_stats,
  by.x = "id",
  by.y = "char_id"
)
```
:::

## Exercise 1

Now, let's make a nice plot out of the data we've got. 


0. First of all, let's clean up our questions column a bit. Replace all "_" characters with "/" characters to make it clearer that we have two poles. Use the interent to figure out which function you need. 

::: {.callout-tip collapse="true"}
## Hint
Use `gsub()`. Look at `?gsub()` to see how it works. 

:::

::: {.callout-caution collapse="true"}
## Solution

```{r}
characters_stats$question <- gsub("_", "/", characters_stats$question)

```
:::

1.  Select up to `40` questions you want to display in the plot and save them into a new vector.

::: {.callout-caution collapse="true"}
## Solution

I'm just gonna take the first 40 questions from all unique ones:

```{r}
questions <- unique(characters_stats$question)[1:40]
questions
```
:::

2.  Select one show (your favorite), extract it from the data frame and save it into a new data set. It should only contain the questions you have selected in the first step.

::: {.callout-caution collapse="true"}
## Solution
```{r}
characters_subset <- characters_stats %>%
  filter(
    uni_name %in% c("Friends"),
    question %in% questions
  )
```
:::

3. Build the coordinate system of the plot. Plot the `rating` on the x axis and the `question` on the y axis.

::: {.callout-caution collapse="true"}
## Solution

::: tidy
```{r}
p <- ggplot(
  data = characters_subset,
  aes(
    x = rating,
    y = question
  )
) 

p
```
:::

:::


4. Now let's split up the plot, so every character gets an own pane. Use `facet_grid()` to do that.

::: {.callout-tip collapse="true"}
## Hint
use `?facet_grid` to find out how to use it. 
:::

::: {.callout-caution collapse="true"}
## Solution

:::tidy
```{r}
p <- p +
  facet_grid(. ~ name)
p
```
:::
:::

5. Let's add bars to the plot by using `geom_col()`. The filling of the bars should depend on the `rating`. You might also want to change the `width` of the bars to fit everything on the page.

::: {.callout-tip collapse="true"}
## Hint
You can change a lot in the appearance of the bars. For example, you might want to use `width` to make the bars a bit smaller. Or use `color` to give them a frame.
:::

::: {.callout-caution collapse="true"}
## Solution

:::tidy
```{r}
p <- p +
  geom_col(
    aes(fill = rating),
    colour = "black",
    width = 0.5
  )

p
```

:::

:::

6. Style the plot. You could choose another color palette, another theme, other labels ... Get creative!

::: {.callout-caution collapse="true"}
## Solution

::: tidy
```{r}

p +
  scale_fill_viridis_c(option = "D") + 
  theme_bw() +
  ggtitle("Characteristics of the 'Friends' characters") +
  xlab("Rating") +
  ylab("")

```

Great! With this color scale we can easily spot if a character is more balanced in his/her personality characteristics (like Chandler Bing), or tends to be pretty extreme (like Monica Geller).
:::

:::
