---
title: "Plotting: Exercises"
---

::: callout-note
These exercises are optional.
:::

::: {.callout-caution icon="false" collapse="true"}
## Previous code

```{r, message = FALSE}
library(tidyverse)

## Load the data
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

## Reshape psych_stats into long format:
psych_stats <- psych_stats %>%
  pivot_longer(cols = messy_neat:innocent_jaded, 
               names_to = "question", 
               values_to = "rating")

## Merge it
characters_stats <- merge(
  x = characters,
  y = psych_stats,
  by.x = "id", 
  by.y = "char_id"
)
```
:::

## Exercise 1

Now, let's make a nice plot out of the data we've got.

1.  Select up to `15` questions you want to display in the plot and save them into a new vector.

::: {.callout-caution collapse="true"}
## Solution

I'm just gonna take the first 15 questions from all unique ones:

```{r}
questions <- unique(characters_stats$question)[1:10]
questions
```
:::

2.  Select one show (your favorite), extract it from the data frame and save it into a new data set. It should only contain the questions you have selected in the first step.

```{r}
characters_subset <- characters_stats %>% 
  filter(uni_name %in% c("Friends"), 
         question %in% questions)
```

```{r eval = FALSE, echo = FALSE}

ggplot(data = characters_subset, 
       aes(x = question, 
           y = rating,
           fill = name)
       ) +
  geom_col(position = "dodge", colour = "black") +
  scale_fill_brewer(palette = "Dark2") +
  theme_bw()
```

```{r eval = FALSE, echo = FALSE}

ggplot(data = characters_subset, 
       aes(x = rating, 
           y = question,
           fill = name)
       ) +
  geom_col(position = "dodge", colour = "black") +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(. ~ name) +
  theme_bw() 

## second axis?
  
```
