---
title: "Plotting"
---

::: {.callout-note, collapse="true"}
## Open if you want to learn more about plotting in R (optional)
This section is for more experienced R users who want to start plotting with `ggplot2`. It is not necessary to follow the rest of the conference.

Now, let's take a closer look at how `ggplot2` works. 
We already had a quick glimpse at it: Plots are build from different layers to create complex outputs. There are endless possibilities for different plot types, [look] here(https://r-graph-gallery.com/) for some inspiration and code. 

First, let's plot a relatively simple plot to get you familiar with how `ggplot2` works. After that, we will the preparation we have done in the last chapters to use and plot the number of gold medals each country has won over the years on a world map, which gets slightly more complex.

Let's load our data:
```{r}
library(tidyverse)

athletes <- readRDS(file = here::here("raw_data", "athletes.rds"))
```


## First plot
For this plot, we want to find the country with the most medal winners for each sport. The preparation is pretty similar to what we have done [before]().

::: tidy
```{r}
best_by_sport <- athletes %>%
  ## Get all gold medalists
  filter(Medal == "Gold") %>%
  ## Group them by sport and region
  group_by(Sport, Region) %>%
  ## count the number of medals each country has per sport category
  count(Medal) %>%
  ## Now only group by sport, so we can extract the maximum medal row by sport, and not by sport and country
  group_by(Sport) %>%
  ## Extract the country with the most medals
  slice(which.max(n)) 
```


### `ggplot()`
In general, a `ggplot` starts with the `ggplot()` function. In it we define the data we want to use, and some aesthetics. The `ggplot()` function then draws our (currently still empty) plotting area, with the defined axes (see next section).

### `aes()`
Aesthetics set parameters **dependent on the data**. In most cases, we will define our x and y axis here. We can also group data together by groups found in a column. If we want the data to have a different color, form, filling etc. depending on values in a column, we can define that here as well (we will look at that later on).

```{r}
#| fig-height: 8
#| fig-width: 11

p <- ggplot(
  data = best_by_sport,
  aes(
    x = Sport,
    y = n,
  )
)

p
```

In this case, the sport is plotted on the x axis and the number of gold medals (`n`) on the y axis. Bar filling will vary depending on the region. 

### geom_()
The geoms do the actual plotting. For example, if we want a barplot:
```{r}
#| fig-height: 8
#| fig-width: 11

p +
  geom_col()
```

Looking pretty boring, right? Let's give each country another color by defining the `fill` aesthetic. Also, lets order the x axis depending on the number of gold medalists:
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  geom_col(aes(fill = Region, x = reorder(Sport, Region)))
p
```
Correct! We can define the aesthetics also within the `geom_()` functions. In this case they will only be used for that specific function, and not for the whole plot (which would be the case if we had defined the `fill` aesthetic in the `ggplot()` function).

```{r}
#| fig-height: 8
#| fig-width: 11

p +
  geom_point()
```

This gives us black points now. If we had defined our aesthetics in the `ggplot()` function, the points would have the same colors as the bars. Also note that we layed another geom (`geom_point()`) over our plot, we added a new layer (that's what I meant earlier on by saying plots consist of different layers). 

### General plotting options
We can tweak all aspects of the appearance of a plot. For example, we might want to turn the x axis labels by 90 degrees to actually make them readable:
```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p
```

Or we could label the bars with the country: 
```{r}
#| fig-height: 8
#| fig-width: 11
 
p <- p +
  geom_text(aes(label = Region), hjust = -0.3, angle = 90, size = 2.5)
p
```

Or use different colors and a different theme:

```{r}
#| fig-height: 8
#| fig-width: 11

p <- p +
  theme_classic() +
  ## And turn the axis labels again, because the new theme has overwritten our theme
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ## Specify which colors are used for the filling
  scale_fill_manual(values = viridisLite::viridis(19))

p  
```

Finally, change the title and axis labels:
```{r}
#| fig-height: 8
#| fig-width: 11

p +
  ggtitle("Country with the most Olympic gold medal winners by sport") +
  xlab("Sport") +
  ylab("Number of gold medal winners")
```

:::

Pretty cool! Now we know that the UK is the all time Olympic Tug-Of-War champion! Also note that we are looking at the number of people from each country winning a gold medal, so team sports are counted multiple times. 

## Second plot
Let's take a look at another, slightly different plot, just to see what can be easily achieved with `ggplot2` (and we have prepared the data to do that, so we don't want to wast the work).

Here is our data preparation from the previous chapters:

```{r}
world_coordinates <- readRDS(file = here::here("raw_data", "world_coordinates.rds"))

medal_counts <- athletes %>%
  filter(Medal == "Gold") %>%
  group_by(Region) %>%
  count(Medal) 

medal_countries <- world_coordinates %>%
  rename("Region" = region) %>%
  left_join(medal_counts)

```


```{r}
#| fig-height: 8
#| fig-width: 11

ggplot(data = medal_countries, 
       mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  geom_polygon(aes(fill = n)) +
  scale_fill_distiller(palette ="YlOrBr", direction = 1) + # or direction=1
  ggtitle("NUmber of Olympic gold medalists by country") +
  theme_void()
```

Note that the medals for team sports are counted multiple times, as our data contained the gold medal wins by person, and not by discipline. 

:::
