---
title: "Loading data: Exercises"
---

## Exercise 1

1.  Download the data set [`characters`](https://github.com/nickhaf/r_tutorial/tree/main/raw_data).

::: {.callout-tip collapse="true"}
## Hint

Right-click on the respective data set found under this link, select 'save as' and then choose a folder on your notebook where to save it. We have talked about where that optimally should be in [Workflow: Exercises](../workflow/worflow_exercise.qmd).
:::

2.  Load it into R, and assign the name `characters`.

::: {.callout-tip collapse="true"}
## Hint

It is a `.rds` file.
:::

::: {.callout-caution collapse="true"}
## Solution

```{r}
characters <- readRDS(file = here::here("raw_data", "characters.rds"))
```
:::

## Exercies 2

::: {.callout-note collapse="true" icon="false"}
## Advanced exercise

1.  Download the [`psych_stats`](https://github.com/nickhaf/r_tutorial/tree/main/raw_data) data set.
2.  Load it into R and assign the name `psych_stats`. Did that work as expected? If not, why not?

::: {.callout-tip collapse="true"}
## Hint

Take a look at the `.csv` file by opening it in a text editor. Look at the documentation of `read.csv()` and take a look at the `sep` argument. Can you define another separator, so R can deduce which elements need to be separated into different cells?
:::

::: {.callout-caution collapse="true"}
## Solution

Let's try the `read.csv()` function:

```{r}
psych_stats <- read.csv(file = here::here("raw_data", "psych_stats.csv"))

head(psych_stats)
```

Huh, that looks weird. If we take a look at the file by opening it in a text editor, we can see that the values are separated by `;`. So let's call the help for `read.csv()`:

```{r, eval = FALSE}
?read.csv
```

The `sep` argument specifies that the seperator needs to be a white space (meaning tabs, spaces ... - look at the details in the documentation).
So, we can do the following:

```{r}
psych_stats <- read.csv(
  file = here::here("raw_data", "psych_stats.csv"),
  sep = ";"
)

str(psych_stats)
```

That looks better!
:::
:::
