
# Exercises

1.  For the rest of the workshop, we will use the data frame `winners` from the package [LondonMarathon](https://github.com/nrennie/LondonMarathon). I have already tweaked it a bit to make it easier to work with, so just download it from [here](https://github.com/nickhaf/r_tutorial/tree/main/raw_data) and load it into R (write the code for reading in the data into a new script).
2.  Advanced: Load the `continents.csv` file into R. Did that work as expected? If not, why not? ([Solution](#Solution)) *Hint*: Take a look at the .csv by opening it in a text editor. Look at the documentation of `read.csv()` and take a look at the `sep` argument.

# Solution {#solution}

```{r}
winners <- readRDS(file = here::here("raw_data", "winners.rds"))
```


```{r, eval = FALSE}
#| code-fold: true
#| code-summary: "Solution for exercise 1"

winners <- readRDS(file = "./raw_data/winners.rds")
```

```{r, eval = FALSE}
#| code-fold: true
#| code-summary: "Solution for exercise 2"
continents <- read.csv(file = "./raw_data/continents.csv")

head(continents)
 
# Hmm that looks weird. If we take a look at the file by opening it in a text editor, we can see that the values are seperated by `;`. 
# So let's call the help for `read.csv()`:

?read.csv

# The `sep` argument specifies that the seperator needs to be a white space (meaning tabs, spaces ... - look at the details).  
# So, we can do the following:

continents <- read.csv(
  file = "./raw_data/continents.csv",
  sep = ";"
)

head(continents)

# This looks better! Now we can merge both data sets, so we can actually work with them.
```

