---
title: "Basic Operations: Exercises"
---

## Exercise 1

What does the function `seq` do?

-   [ ] Repeats a value multiple times.
-   [ ] Builds a sequence of values.
-   [ ] Loads a SQL data base.
-   [ ] It's part of another package and therefore not loaded in `Base R`.

::: {.callout-tip collapse="true"}
## Hint

Use the help function `?`.
:::

::: {.callout-caution collapse="true"}
### Solution

-   [ ] Repeats a value multiple times.
-   [x] Builds a sequence of values.
-   [ ] Loads a SQL data base.
-   [ ] It's part of another package and therefore not loaded in R.
:::

## Exercise 2

Why does the following code not work? Correct it so it does.

```{r assign, eval = FALSE}
c(1, 2, 3, 4, 5)
mean(num_vec)
```

::: {.callout-tip collapse="true"}
## Hint

Does the object `num_vec` actually exist?
:::

::: {.callout-caution collapse="true"}
## Solution

The object `num_vec` hasn't been assigned yet. So let's do that:

```{r}
num_vec <- c(1, 2, 3, 4, 5)
mean(num_vec)
```
:::

## Exercise 3

Build the following vector with as little code as possible:

```{r build_vec, eval = FALSE}
vec_1 <- c(1.0, 2.0, 3.0, 4.0, 5.0, 5.1, 5.2, 5.3, 5.4, 5.5, 2.0, 2.0, 2.0, 2.0, 2.0, 2.0, 2.0, 2.0)
```

::: {.callout-tip collapse="true"}
## Hint

Use `seq()` and `rep()`. You can also build consecutive sequences using `:`.
:::

::: {.callout-caution collapse="true"}
## Solution

```{r build_vec-solution}
vec_1 <- c(1:5, seq(from = 5, to = 5.5, by = 0.1), rep(x = 2, times = 8))
vec_1
```
:::

## Exercise 4

Find all of the elements in the vector `vec_num` that are either equal to `1000`, or lie between `sqrt(11)` and `log(1.001)`.

```{r log}
vec_num <- c(sqrt(100)^3, exp(-6), 22.02/3 * sqrt(4^2) * 0.25, -120987/(47621 * 1.3 ^ 4 ))
```

::: {.callout-tip collapse="true"}
## Hint

You need to combine three logical statements. Go at it step by step: first find all elements in `vec_num` that are equal to `1000`, and then add a comparison for the rest of the statement behind an `|` ("or").
:::

::: {.callout-caution collapse="true"}
## Solution

```{r log-solution}
vec_num == 1000 | (vec_num < sqrt(11) & vec_num > log(1.001))
```
:::
