---
title: "Final Exercise"
---

[^1]: Image by [Mitchell Luo](https://unsplash.com/de/@mitchel3uo?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText) on [Unsplash](https://unsplash.com/de/fotos/2Oazk1nJzhM?utm_source=unsplash&utm_medium=referral&utm_content=creditCopyText).


This exercise comes back to most topics presented in this workshop. If you are a R beginner and followed the workshop, you can do this last exercise in the end to again test your knowledge. If you already have some R experience, you can do this exercise before the rest of the workshop and use it to identify weak points to follow up on. 

So, in order to provide you with a totally fresh data set, let's look at beach volleyball, even though the others might be worth further exploring as well. 

## Exercise 1
1. [Download](../qmd/load_data/load_data_exercise.qmd) the data set [`volleyball`](https://github.com/nickhaf/r_tutorial/tree/main/raw_data) and load it into R. 

::: {.callout=caution collapse="true"}
## Solution
```{r}
vb <- tidytuesdayR::tt_load('2020-05-19')$vb_matches

```
:::


2. Our goal will be to compare the mean height of the winning team against the mean height of the losing team. Select only the columns from the data frame that are relevant to this question 


::: {.callout=caution collapse="true"}
## Solution

```{r}
vb <- vb[ , c()]
```

::: 


3. Merge losing on winning team.


4. Remove NAs.

5. Look at distribution.


::: {.callout=caution collapse="true"}
Huh that is weird? Height should be normally distributed, not bimodal. 
Ah, we still have men and women in the same data set. 
So group by gender.

Ah, that looks better. 
 
:::

6. Calculate the mean height per team. 

7. Build two separate data sets, one for men, one for women. 


::: {.callout=note collapse="true"}
We haven't really looked at most of this stuff in depth. Use your favorite search engine and try to solve it anyways, as you will often encounter problems you haven't seen yet. 

8. Do a t test comparing the mean winners height against the mean losers height. Use a loop to iterate over both data sets, so you can calculate the t test for men and women separately. Save the results in a list.

9. Look at the mean difference. Not very big in most cases, right? Let's calculate a standardized effect size, Cohen's *d*:


Write a function to do that, and add the function to your for loop, so it automatically returns the effect size along with the p value. 
:::

10. For plotting, add the rows of the men's data frame to the bottom of the women's data frame.


11. Add a new column containing the height in meters, not in Inch (one Inch are 0.0254 m). 

12. Reform this resulting data frame into long format.

13. Use ggplot to build a violin plot showing the winners and losers distribution by gender. Violin plots are similar to boxplotsy but have the advantage of conveying more distributional information.


```
tuesdata$winning_height <- rowMeans(tuesdata[, c("w_p1_hgt", "w_p2_hgt")]) 
tuesdata$losing_height <- rowMeans(tuesdata[, c("l_p1_hgt", "l_p2_hgt")]) 

volleyball_data_men <- tuesdata %>% 
  filter(gender == "M")


hist(volleyball_data_men$winning_height)
hist(volleyball_data_men$losing_height)

t.test(volleyball_data_men$winning_height, volleyball_data_men$losing_height)

## Transform into long for plotting


## Plot boxplot for winner and loser distribution

ggplot(volleyball_data_men) +
  geom_boxplot(aes(y = ))

## Zusatz:
## Write a function to calculate cohens d


## Write a function for inch into meter:
inch_m <- function(inch){
  m <- inch * 0.0254
  return(m)
}

volleyball_data_men$winning_height <- inch_m(volleyball_data_men$winning_height)
volleyball_data_men$losing_height <- inch_m(volleyball_data_men$losing_height)


```

